FTLR proof cases

### 2018-03-14 fully-abstract-comp/ftlr-proof-cases.tex

Let $n' \leq n$ be given and assume
\begin{enumerate}
\item $\npair[n']{[(c,c)]} \in \lrp(W)$ 
\dominique{Weird that this is an assumption here.  Shouldn't this be a proof obligation? Or rather: I think this should not be here at all but in the value relation instead? Remove this from the definition of $\lre$?}
\lau{$(c,c)$ is not shown or assumed to be in the value relation, so to make sure it respects lineraity, this condition is placed here. Further, this does become a proof obligation when you use the FTLR.}
\item $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_R)$ \label{item:reg-ass}
\item $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$\label{item:mem-ass}
\item $W \oplus W_R \oplus W_M$ is well-defined\label{item:worlds-def}
\end{enumerate}
Further let
\begin{itemize}
\item $\Phi_S = (\ms_S,\reg_S\updReg{\pcreg}{c},\stk,\ms_\stk)$
\item $\Phi_T = (\ms_T,\reg_T\updReg{\pcreg}{c})$
\end{itemize}
and show
\[
  \npair[n']{(\Phi_S,\Phi_T)} \in \lro
\]
Let $i \leq n'$ be given \footnote{ignoring the frame...} and show
\[
  \Phi_S \term[i] \Rightarrow \Phi_T \term
\]
and 
\[
  \Phi_T \term[i] \Rightarrow \Phi_S \term
\]

@@

In case $\scall[o]{}{r_1}{r_2}$ is the first step of the execution. We can assume $r_1,r_2 \neq \pcreg$ as this would cause the execution to fail\footnote{Assuming $\pcreg$ was one of the two registers, then the step relation would reqiure the register to contain an executable capability while the call semantics would require there to be a sealed capability.}\footnote{{\color{red} TODO:} We still have to consider what happens if the temp register is jumped to.}. By assumption, $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_R)$ which in particular means that for suitable $W_{R,1}$ and $W_{R,2}$

@@

  (\Phi_S(r_2), \vsc_{T,2}) \in H_\sigma \; \sigma \; W_{R,2}

@@

At this point, we need to pick a suitable $W' \future W_{R,1}$. To this end, we need to know a bit about $W_{R,1}$. From the operational semantics, we know $\Phi_S(\rstk) = \stkptr{\perm,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}$ and by assumption $\Phi_S(\rstk) \in \lrv(W_{R,\stk})$. From this we can conclude that there is a stack region (specifically a spatially owned region) $r$ for which $W_{R,\stk} \nequal[n'-1] \stdreg{[\baddr_\stk',\eaddr_\stk']}{\spa}$ where $[\baddr_\stk',\eaddr_\stk'] \supseteq [\baddr_\stk,\eaddr_\stk]$\footnote{Actually concluding this involves knowing that $W_{R,\stk}$ is ``well-formed'' which we use the heap satisfaction assumption to conclude. Perhaps a notion of well-formed worlds would be good along with a couple of lemmas that use it to make this reasoning once and for all.}. As $W_{R,\stk} \oplus W_{R,1}$ is well-formed, $W_{R,1}(r)$ must be a similar spatial region to the one in $W_{R,\stk}$.

@@

               & \update{\mathrm{priv}.y}{(\stdreg{[\aaddr_\stk,\eaddr_\stk']}{\spa},\opc)}

@@

where $x$ and $y$ are fresh region names in the respective part of the world. To indicate similar worlds to $W'$ with the same ownership as another world well-defined w.r.t. $W$, we write $W'_{R,1}$ with the same spatially owned region as $W_{R,1}$. Using the private part of the future world relation, it is easy to see that $W'$ is a future world of $W_{R,1}$ (by the operational semantics $\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk$). Using world monotonicity of $H_\sigma \; \sigma$\footnote{{\color{red}This means that $H_\sigma$ must be monotone in the world!}}, we get

@@

  \npair[n'']{\Phi_S(r_1),\Phi_S(r_2),\Phi_T(r_1),\Phi_T(r_2)} \in \lrexj(W_{R,1}' \oplus W_{R,2}')

@@

For the target side, we pick the register-file
\[
  \Phi_T.\reg
  \begin{array}[t]{l}
    \update{r_1}{\linCons{\Phi_T(r_1)}}\\
    \update{r_2}{\linCons{\Phi_T(r_2)}}\\
    \update{\pcreg}{0} \\
    \update{\rdata}{0} \\
    \update{\rstk}{c_\stk'}\\
    \update{\rretc}{c_{\ret,c}'}\\
    \update{\rretd}{c_{\ret,d}'}\\
    \update{\rtmp{1}}{0}\\
  \end{array}  
\]
where
\begin{itemize}
\item $c_\stk' = ((\rw,\linear),\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1)$
\item $c_{\ret,c}' = \sealed{\sigma',((\perm,\lin),\baddr,\eaddr,\aaddr+\calllen-\dots)}$
\item $c_{\ret,d}' = \sealed{\sigma',((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk-1)}$
\end{itemize}
\lau{We do not actually check the linearity of the stack, but it should have been checked that the base address actually is the stack base address which would give us that it is linear. This is not an assumption we have formally stated at the moment!}
and the same memory
\[
  \Phi_T.\ms_T\update{\aaddr_\stk}{42}
\]
We have to argue (1) that this register-file pair is in the register-file relation for a world $W_R'$ with the same ownership as $W_R$ and regions $z$ and $y$ are owned, and (2) the memory-pair satisfies the memory satisfaction relation for a world $W_M'$ (i.e. with the same ownership as $W_M$).

Let's start with (2). Here we use that we took a standard region and replaced by two standard regions. This means that when we split up the memory, we can reuse the same splitting og memory and world (w.r.t memory) except for the memory segment that previously satsfied the standard region in question. This memory segment, we split in two, so each piece fit with one of the new regions.

@@

For $\rstk$, we make sure the $z$ region is owned in the world we use. Now we use that we had a standard region governing the stack to get that everything on there is in the value relation. We then use the monotonicity of the value relation to conclude that everything is in the value relation in this future world which satisfies the read and write condition.

>>>>

Let $n' \leq n$ be given and assume
\begin{enumerate}
\item $\npair[n']{[(c,c)]} \in \lrp(W)$ 
\dominique{Weird that this is an assumption here.  Shouldn't this be a proof obligation? Or rather: I think this should not be here at all but in the value relation instead? Remove this from the definition of $\lre$?}
\lau{$(c,c)$ is not shown or assumed to be in the value relation, so to make sure it respects lineraity, this condition is placed here. Further, this does become a proof obligation when you use the FTLR.}
\item $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_R)$ \label{item:reg-ass}
\item $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$\label{item:mem-ass}
\item $W \oplus W_R \oplus W_M$ is well-defined\label{item:worlds-def}
\end{enumerate}
Further let
\begin{itemize}
\item $\Phi_S = (\ms_S,\reg_S\updReg{\pcreg}{c},\stk,\ms_\stk)$
\item $\Phi_T = (\ms_T,\reg_T\updReg{\pcreg}{c})$
\end{itemize}
and show
\[
  \npair[n']{(\Phi_S,\Phi_T)} \in \lro
\]
That is, let $i \leq n'$ be given \footnote{ignoring the frame...} and show
\[
  \Phi_S \term[i] \Rightarrow \Phi_T \term
\]
and 
\[
  \Phi_T \term[i] \Rightarrow \Phi_S \term
\]

@@

In case $\scall[o]{}{r_1}{r_2}$ is the first step of the execution. We can assume $r_1,r_2 \neq \pcreg$ as this would cause the execution to fail\footnote{Assuming $\pcreg$ was one of the two registers, then the step relation would reqiure the register to contain an executable capability while the call semantics would require there to be a sealed capability.}\footnote{{\color{red} TODO:} We still have to consider what happens if the temp register is jumped to.}. By assumption \ref{item:reg-ass}, $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_R)$ which in particular means that for suitable $W_{R,1}$ and $W_{R,2}$

@@

  (\Phi_S(r_2), \vsc_{T,2}) \in H_\sigma \; \sigma \; W_{R,2}.

@@

At this point, we need to pick a suitable $W' \future W_{R,1}$. To this end, we need to know a bit about $W_{R,1}$. From the operational semantics and in particular the semantics of $\scall{}{}{}{}$, we know $\Phi_S(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}$ and by assumption $\Phi_S(\rstk) \in \lrv(W_{R,\stk})$. From this we can conclude that there is a stack region (specifically a spatially owned region) $r$ for which $\pwfree{W_{R,\stk}}(r) \nequal[n'-1] \stdreg{[\baddr_\stk',\eaddr_\stk']}{\spa}$ where $[\baddr_\stk',\eaddr_\stk'] \supseteq [\baddr_\stk,\eaddr_\stk]$\footnote{Actually concluding this involves knowing that $W_{R,\stk}$ is ``well-formed'' which we use the heap satisfaction assumption to conclude. Perhaps a notion of well-formed worlds would be good along with a couple of lemmas that use it to make this reasoning once and for all.}. As $W_{R,\stk} \oplus W_{R,1}$ is well-formed, $\pwfree{W_{R,1}}(r)$ must be a similar spatial region to the one in $W_{R,\stk}$.

@@

               & \update{\mathrm{priv}.y}{(\stdreg{[\aaddr_\stk,\eaddr_\stk]}{\spa},\opc)}\\
               & \update{\mathrm{free}.z}{\stdreg{[\eaddr_\stk+1,\addr_\stk']}{\spa}}

@@

where $x$, $y$, and $z$ are fresh region names in the respective part of the world. To indicate similar worlds to $W'$ with the same ownership as another world well-defined w.r.t. $W$, we write $W'_{R,1}$ with the same spatially owned region as $W_{R,1}$. Using the private part of the future world relation, it is easy to see that $W'$ is a future world of $W_{R,1}$ (by the operational semantics $\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk$, so the new region at $y$ is non-empty). Using world monotonicity of $H_\sigma \; \sigma$\footnote{{\color{red}This means that $H_\sigma$ must be monotone in the world!}}, we get

@@

  \npair[n'']{(\Phi_S(r_1),\Phi_S(r_2),\Phi_T(r_1),\Phi_T(r_2))} \in \lrexj(W_{R,1}' \oplus W_{R,2}')

@@

For the target side, we pick the register-file
\[
  \Phi_T.\reg
  \begin{array}[t]{l}
    \update{r_1}{\linCons{\Phi_T(r_1)}}\\
    \update{r_2}{\linCons{\Phi_T(r_2)}}\\
    \update{\pcreg}{0} \\
    \update{\rdata}{0} \\
    \update{\rstk}{c_\stk'}\\
    \update{\rretc}{c_{\ret,c}'}\\
    \update{\rretd}{c_{\ret,d}'}\\
    \update{\rtmp{1}}{0}\\
  \end{array}  
\]
where
\begin{itemize}
\item $c_\stk' = ((\rw,\linear),\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1)$
\item $c_{\ret,c}' = \sealed{\sigma',((\perm,\lin),\baddr,\eaddr,\aaddr+\calllen-\dots)}$
\item $c_{\ret,d}' = \sealed{\sigma',((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk-1)}$
\end{itemize}
\lau{We do not actually check the linearity of the stack, but it should have been checked that the base address actually is the stack base address which would give us that it is linear. This is not an assumption we have formally stated at the moment!}
and the same memory
\[
  \Phi_T.\ms_T\update{\aaddr_\stk}{42}
\]
We have to argue (1) that this register-file pair is in the register-file relation for a world $W_R'$ with the same ownership as $W_R$ and regions $x$ and $y$ are owned, and (2) the memory-pair satisfies the memory satisfaction relation for a world $W_M'$ (i.e. with the same ownership as $W_M$).

Let's start with (2). Here we use that we took a standard region and replaced by three standard regions. This means that when we split up the memory, we can reuse the same splitting of memory and world (w.r.t memory) except for the memory segment that previously satsfied the standard region in question. This memory segment, we split in three, so each piece fit with one of the new regions.

@@

For $\rstk$, we make sure the $x$ region is owned in the world we use. Now we use that we had a standard region governing the stack to get that everything on there is in the value relation. We then use the monotonicity of the value relation to conclude that everything is in the value relation in this future world which satisfies the read and write condition.
